Unicode Implementation Issues \thanks { This is a working document for the Rexx Language Association (RexxLA). There will be one or more separate requirements documents. It is written in \LaTeX\ \fmtversion\ and was rendered \today\ using \enginedetails. }

### 2024-04-12 ARB/standards/work-in-progress/document/tex/standard/configuration.tex
add filenames

We don't want to add undue ``magic'' to this section. It seems we will
need the concept of a ``reference'' (equivalent to a machine address) so
that this section can at least have composite objects as arguments. (As
it already does but these are not Rexx objects)

Possibly we could unify ``reference'' with ``variable pool number''
since object one-to-one with its variable pool is a fair model. That way
we don't need a new primitive for comparison of two references.

JAVA is only a ``reference'' for NetRexx so some generalized JAVA-like
support is needed for that. It would provide the answers to what classes
were in the context, what their method signatures were etc.

@@

\begin{itemize}
\item
  Config\_ is used only from the language processor when processing
  programs;
\item
  API\_is part of the application programming interface and is
  accessible from programs which are not written in the Rexx language;
\item
  Trap\_ is not provided by the language processor but may be invoked by
  the language processor. As its result, each function shall return a
  completion Response. This is a string indicating how the function
  behaved. The completion response may be the character `N' indicating
  the normal behavior occurred; otherwise the first character is an
  indicator of a different behavior and the remainder shall be suitable
  as a human-readable description of the function's behavior.
\end{itemize}

This standard defines any additional results from Config\_ functions as
made available to the language processor in variables. This does not
constrain how a particular implementation should return these results.

@@

As alternatives to the normal indicator `N', each function may return a
completion response with indicator 'X' or `S'; other possible indicators
are described for each function explicitly. The indicator `X' means that
the function failed because resources were exhausted. The indicator `S'
shows that the configuration was unable to perform the function. Certain
indicators cause conditions to be raised. The possible raising of these
conditions is implicit in the use of the function; it is not shown
explicitly when the functions are used in this standard. The implicit
action is call \#Raise `SYNTAX', Message, Description where: \#Raise
raises the condition, see nnn. Message is determined by the indicator in
the completion response. If the indicator is `X' then Message is 5.1. If
the indicator is `S' then Message is 48.1. Description is the
description in the completion response. The `SYNTAX' condition 5.1 can
also be raised by any other activity of the language processor.

@@

The processing initiation interface consists of a function which the
configuration shall provide to invoke the language processor. We could
do REQUIRES in a macro-expansion way by adding an argument to
Contig\_SourceChar to specify the source file. However, I'm assuming we
will prefer to recursively ``run'' each required file. One of the
results of that will be the classes and methods made public by that
REQUIRES subject.

@@

API Start(How, Source, Environment, Arguments, Streams, Traps, Provides)

where: How is one of `COMMAND', `FUNCTION', or `SUBROUTINE' and
indicates how the program is invoked.

What does OOI say for How when running REQUIREd files?

Source is an identification of the source of the program to be
processed.

Environment is the initial value of the environment to be used in
processing commands. This has components for the name of the environment
and how the input and output of commands is to be directed.

Arguments is the initial argument list to be used in processing. This
has components to specify the number of arguments, which arguments are
omitted, and the values of arguments that are not omitted.

Streams has components for the default input stream to be used and the
default output streams to be used.

Traps is the list of traps to be used in processing (see nnn). This has
components to specify whether each trap is omitted or not.

Semantics:

@@

If the program was terminated due to a RETURN or EXIT instruction
without an expression the completion response is `N'.

If the program was terminated due to a RETURN or EXIT instruction with
an expression the indicator in the completion response is `R' and the
description of the completion response is the value of the expression.

If the program was terminated due to an error the indicator in the
completion response is `E' and the description in the completion
response comprises information about the error that terminated
processing.

If How was `REQUIRED' and the completion response was not `E', the
Provides argument is set to reference classes made available. See nnn
for the semantics of these classes.

@@

\begin{itemize}
\item
  syntactic\_characters;
\item
  extra\_letters;
\item
  other\_blank\_characters;
\item
  other\_negators;
\item
  other\_characters.
\end{itemize}

@@

The following characters represent the category of characters called
syntactic\_characters, identified by their names. The glyphs used to
represent them in this document are also shown. Syntactic\_characters
shall be available in every configuration:

\begin{itemize}
\item
  \& ampersand;
\item
  apostrophe, single quotation mark, single quote;
\item
  asterisk, star;
\item
  blank, space;
\item
  A-Z capital letters A through Z;
\item
  colon;
\item
  , comma;
\item
  0-9 digits zero through nine;
\item
  = equal sign;
\item
  exclamation point, exclamation mark;
\item
  \begin{quote}
  greater-than sign;
  \end{quote}
\end{itemize}

hyphen, minus sign;

\textless{} less-than sign;

\begin{itemize}
\item
  {[} left bracket, left square bracket; (\_ left parenthesis;
\item
  \% percent sign;
\item
  . period, decimal point, full stop, dot;
\item
  plus sign;
\item
  ? question mark;
\item
  '' quotation mark, double quote; ~reverse slant, reverse solidus,
  backslash; {]} right bracket, right square bracket;
\item
  ) right parenthesis; ; semicolon; /\_ slant, solidus, slash; a-z small
  letters a through z;
\item
  \textasciitilde{} tilde, twiddle;
\item
  \_ underline, low line, underscore;
\item
  vertical line, bar, vertical bar.
\end{itemize}

@@

A configuration may have a category of characters in source programs
called extra\_letters. Extra\_letters are determined by the
configuration.

@@

A configuration may have a category of characters in source programs
called other\_blank\_characters. Other\_blank\_characters are determined
by the configuration. Only the following characters represent possible
characters of this category:

\begin{itemize}
\item
  carriage return;
\item
  form feed;
\item
  horizontal tabulation;
\item
  new line;
\item
  vertical tabulation.
\end{itemize}

@@

A configuration may have a category of characters in source programs
called other\_negators. Other\_negators are determined by the
configuration. Only the following characters represent possible
characters of this category. The glyphs used to represent them in this
document are also shown:

\begin{itemize}
\tightlist
\item
  \begin{itemize}
  \tightlist
  \item
    circumflex accent, caret;
  \end{itemize}
\item
  --- not sign.
\end{itemize}

@@

A configuration may have a category of characters in source programs
called other\_characters. Other\_characters are determined by the
configuration.

@@

The configuration characters and encoding interface consists of
functions which the configuration shall provide which are concerned with
the encoding of characters. The following functions shall be provided:

\begin{itemize}
\item
  Config\_SourceChar;
\item
  Config\_OtherBlankCharacters;
\item
  Config\_Upper;
\item
  Config\_Compare;
\item
  Config\_B2C;
\item
  Config\_C2B;
\item
  Config\_Substr;
\item
  Config\_Length;
\item
  Config\_Xrange.
\end{itemize}

@@

Config SourceChar ()

Semantics: Supply the characters of the source program in sequence,
together with the EOL and EOS events. The EOL event represents the end
of a line. The EOS event represents the end of the source program. The
EOS event must only occur immediately after an EOL event. Either a
character or an event is supplied on each invocation, by setting
\#Outcome.

If this function is unable to supply a character because the source
program encoding is incorrect the indicator of the completion response
is `E' and the description of the completion information is the encoding
which is incorrect, in hexadecimal notation.

@@

Syntax: Config OtherBlankCharacters ()

Semantics: Get other\_blank\_characters (see nnn). Set \#Outcome to a
string of zero or more distinct characters in arbitrary order. Each
character is one that the configuration considers equivalent to the
character Blank for the purposes of parsing.

@@

Syntax: Config Upper (Character)

where: Character is the character to be translated to uppercase.
Semantics: Translate Character to uppercase. Set \#Outcome to the
translated character. Characters which have been subject to this
translation are referred to as being in uppercase. Config\_Upper applied
to a character in uppercase must not change the character.

@@

Config Lower (Character) where: Character is the character to be
translated to lowercase. Semantics: Translate Character to lowercase.
Set \#Outcome to the translated character. Characters which have been
subject to this translation are referred to as being in lowercase.
Config\_Lower applied to a character in lowercase must not change the
character. Config\_Upper of the outcome of Config\_Lower(Character)
shall be the original character.

@@

Syntax: Config Compare(Characterl, Character2)

@@

Character1 is the character to be compared with Character2. Character2
is the character to be compared with Character1.

Semantics:

@@

\begin{itemize}
\item
  `equal' if Character1 is equal to Character2;
\item
  `greater' if Character1 is greater than Character2;
\item
  `lesser' if Character' is less than Character2. The function shall
  exhibit the following characteristics. If Config \_Compare(a,b)
  produces
\item
  `equal' then Config\_Compare(b,a) produces `equal';
\item
  `greater' then Config\_Compare(b,a) produces `lesser';
\item
  `lesser' then Config\_Compare(b,a) produces `greater';
\item
  `equal' and Config\_Compare(b,c) produces `equal' then
  Config\_Compare(a,c) produces `equal';
\item
  `greater' and Config\_Compare(b,c) produces `greater' then Config
  \_Compare(a,c) produces `greater';
\item
  `lesser' and Config \_Compare(b,c) produces `lesser' then
  Config\_Compare(a,c) produces `lesser';
\item
  `equal' then Config\_Compare(a,c) and Config\_Compare(b,c) produce the
  same value. Syntactic characters which are different characters shall
  not compare equal by Config\_Compare, see nnn.
\end{itemize}

@@

Syntax: Config B2C (Binary) where: Binary is a sequence of digits, each
`0' or `1'. The number of digits shall be a multiple of eight.
Semantics:

@@

Syntax: Config C2B (String)

where: String is a string. Semantics: Translate String to a sequence of
digits, each `0' or `1'. Set \#Outcome to the result. This function is
the inverse of Config\_B2C.

@@

Syntax: Config Substr(String, n)

where: String is a string. nis an integer identifying a position within
String. Semantics: Copy the n-th character from String. The leftmost
character is the first character. Set Outcome to the resulting
character. If this function is unable to supply a character because
there is no n-th character in String the indicator of the completion
response is `M'. If this function is unable to supply a character
because the encoding of String is incorrect the indicator of the
completion response is `E' and the description of the completion
information is the encoding which is incorrect, in hexadecimal notation.

@@

Syntax: Config Length (String)

where: String is a string. Semantics: Set \#Outcome to the length of the
string, that is, the number of characters in the string. If this
function is unable to determine a length because the encoding of String
is incorrect, the indicator of the completion response is `E' and the
description of the completion information is the encoding which is
incorrect, in hexadecimal notation.

@@

Config Xrange(Characterl, Character2) where: Character1 is the null
string, or a single character. Character2 is the null string, or a
single character. Semantics: If Character1 is the null string then let
LowBound be a lowest ranked character in the character set according to
the ranking order provided by Config\_Compare; otherwise let LowBound be
Character1. If Character2 is the null string then let HighBound be a
highest ranked character in the character set according to the ranking
order provided by Config\_Compare; otherwise let HighBound be
Character2lf \#Outcome after Config\_Compare(LowBound,HighBound) has a
value of

@@

Syntax: Config ObjectNew

Semantics:

@@

Syntax: Config Array Size(Object, size) where: Object is an object. Size
is an integer greater or equal to 0. Semantics: The configuration should
prepare to deal efficiently with the object as an array with indexes
having values up to the value of size.

@@

Syntax: Config Array Put(Array, Item, Index)

where: Array is an array. Item is an object Index is an integer greater
or equal to 1. Semantics: The configuration shall record that the array
has Item associated with Index.

@@

Syntax: Config Array At(Array, Index)

where: Array is an array. Index is an integer greater or equal to 1.
Semantics: The configuration shall return the item that the array has
associated with Index.

@@

Syntax: Config Array At(Array, Index) where: Array is an array. Index is
an integer greater or equal to 1. Semantics: Return `1' if there is an
item in Array associated with Index, `0' otherwise.

@@

Syntax: Config Array At(Array, Index)

where: Array is an array. Index is an integer greater or equal to 1.
Semantics: After this operation, no item is associated with the Index in
the Array.

@@

Config Command(Environment, Command) where: Environment is the
environment to be addressed. It has components for:

\begin{itemize}
\item
  the name of the environment;
\item
  the name of a stream from which the command will read its input. The
  null string indicates use of the default input stream;
\item
  the name of a stream onto which the command will write its output. The
  null string indicates use of the default output stream. There is an
  indication of whether writing is to APPEND or REPLACE;
\item
  the name of a stream onto which the command will write its error
  output. The null string indicates use of the default error output
  stream. There is an indication of whether writing is to APPEND or
  REPLACE. Command is the command to be executed. Semantics: Perform a
  command.
\item
  set the indicator to `E' or `F' if the command ended with an ERROR
  condition, or a FAILURE condition, respectively;
\item
  set \#RC to the return code string of the command.
\end{itemize}

@@

The external routines interface consists of a function which the
configuration shall provide to invoke external routines. See nnn and nnn
for a description of the language features that use external routines.

@@

Syntax: Config ExternalRoutine(How, NameType, Name, Environment,
Arguments, Streams, Traps) where: How is one of `FUNCTION' or
`SUBROUTINE' and indicates how the external routine is to be invoked.

@@

Semantics:

@@

OOI has external classes explicitly via the ::CLASS abc EXTERNAL
mechanism. Analogy with classic would also allow the subject of
::REQUIRES to be coded in non-Rexx. However ::REQUIRES subject is coded,
we need to gather in knowledge of its method names because of the search
algorithm that determines which method is called. Hence reasonable that
the ultimate external call is to a method. Perhaps combine
Config\_ExternalRoutine with Config\_ExternalMethod. There is a
terminology clash on ``environment''. Perhaps easiest to change the
classic to ``address\_environment''. (And make it part of new
``environment''?) There are terminology decisions to make about
``files'', ``programs'', and ``packages''. Possibly ``program'' is the
thing you run (and we don't say what it means physically), ``file'' is a
unit of scope (ROUTINEs in current file before those in REQUIREd), and
``package'' we don't use (since a software package from a shop would
probably have several files but not everything to run a program.) Using
``file'' this way may not be too bad since we used ``stream'' rather
than ``tile'' in the classic definition. The How parameter will need
`METHOD' as a value. Should API\_Start also allow 'METHOD''. If we pass
the new Environment we don't have to pass Streams separately.

Text of Config\_ExternalMethod waiting on such decisions. Syntax:

Config ExternalMethod (How, NameType, Name, Environment, Arguments,
Streams, Traps) where: How is one of `FUNCTION' or `SUBROUTINE' and
indicates how the external routine is to be invoked. NameType is a
specification of whether the name was provided as a symbol or as a
string literal. Name is the name of the routine to be invoked.
Environment is an environment value with the same components as on API\_
Start. Arguments is a specification of the arguments to the routine,
with the same components as on API\_Start. Streams is a specification of
the default streams, with the same components as on API\_ Start. Traps
is the list of traps to be used in processing, with the same components
as on API\_ Start. Semantics: Invoke an external routine. Set \{Outcome
to the result of the external routine, or set the indicator of the
completion response to `D' if the external routine did not provide a
result. If this function is unable to locate the routine the indicator
of the completion response is `U'. As a result SYNTAX condition 43.1 is
raised implicitly. If How indicated that a result from the routine was
required but the routine did not provide one the indicator of the
completion response is `H'. As a result SYNTAX condition 44.1 is raised
implicitly. If How indicated that a result from the routine was required
but the routine provided one that was too long (see \#Limit\_String in
nnn) the indicator of the completion response is `L'. As a result SYNTAX
condition 52 is raised implicitly. If the routine failed in a way not
indicated by some other indicator the indicator of the completion
response is `F'. As a result SYNTAX condition 40.1 is raised implicitly.
5.8 External data queue The external data queue interface consists of
functions which the configuration shall provide to manipulate an
external data queue mechanism. See nnn, nnn, nnn, nnn, and nnn for a
description of language features that use the external data queue. The
configuration shall provide an external data queue mechanism. The
following functions shall be provided:

@@

Syntax: Config Push(String)

where: String is the value to be retained in the external data queue.
Semantics: Add String as an item to the end of the external data queue
from which Config\_Pull will remove an item.

@@

Syntax: Config Queue (String)

where: String is the value to be retained in the external data queue.
Semantics: Add String as an item to the opposite end of the external
data queue from which Config\_Pull will remove an item.

@@

Syntax: Config Pull()

Semantics: Retrieve an item from the end of the external data queue to
which Config\_Push adds an element to the list. Set \#Outcome to the
value of the retrieved item. If no item could be retrieved the indicator
of the completion response is `F'.

@@

Syntax: Config Queued ()

Semantics:

@@

Config Stream Unique () Semantics: Set \#Outcome to a name that the
configuration recognizes as a stream name. The name shall not be a name
that the configuration associates with any existing data.

>>>>

\emph{We don't want to add undue ``magic'' to this section. It seems we
will need the concept of a ``reference'' (equivalent to a machine
address) so that this section can at least have composite objects as
arguments. (As it already does but these are not Rexx objects)}

\emph{Possibly we could unify ``reference'' with ``variable pool
number'' since object one-to-one with its variable pool is a fair model.
That way we don't need a new primitive for comparison of two
references.}

\emph{JAVA is only a ``reference'' for NetRexx so some generalized
JAVA-like support is needed for that. It would provide the answers to
what classes were in the context, what their method signatures were
etc.}

@@

\begin{itemize}
\item
  \texttt{Config\_} is used only from the language processor when
  processing programs;
\item
  \texttt{API\_} is part of the application programming interface and is
  accessible from programs which are not written in the Rexx language;
\item
  \texttt{Trap\_} is not provided by the language processor but may be
  invoked by the language processor.
\end{itemize}

As its result, each function shall return a completion
\texttt{Response}. This is a string indicating how the function behaved.
The completion response may be the character
\texttt{\textquotesingle{}N\textquotesingle{}} indicating the normal
behavior occurred; otherwise the first character is an indicator of a
different behavior and the remainder shall be suitable as a
human-readable description of the function's behavior.

This standard defines any additional results from \texttt{Config\_}
functions as made available to the language processor in variables. This
does not constrain how a particular implementation should return these
results.

@@

As alternatives to the normal indicator
\texttt{\textquotesingle{}N\textquotesingle{}}, each function may return
a completion response with indicator
\texttt{\textquotesingle{}X\textquotesingle{}} or
\texttt{\textquotesingle{}S\textquotesingle{}}; other possible
indicators are described for each function explicitly. The indicator
\texttt{\textquotesingle{}X\textquotesingle{}} means that the function
failed because resources were exhausted. The indicator
\texttt{\textquotesingle{}S\textquotesingle{}} shows that the
configuration was unable to perform the function.

Certain indicators cause conditions to be raised. The possible raising
of these conditions is implicit in the use of the function; it is not
shown explicitly when the functions are used in this standard.

The implicit action is

\lstinputlisting[language=rexx,label=raisesyntax.rexx,caption=raisesyntax.rexx]{raisesyntax.rexx}

\begin{itemize}
\tightlist
\item
  \texttt{\#Raise} raises the condition, see nnn.
\item
  \texttt{Message} is determined by the indicator in the completion
  response. If the indicator is
  \texttt{\textquotesingle{}X\textquotesingle{}} then \texttt{Message}
  is \texttt{5.1}. If the indicator is
  \texttt{\textquotesingle{}S\textquotesingle{}} then Message is
  \texttt{48.1}.
\item
  \texttt{Description} is the description in the completion response.
\end{itemize}

The \texttt{\textquotesingle{}SYNTAX\textquotesingle{}} condition
\texttt{5.1} can also be raised by any other activity of the language
processor.

@@

The processing initiation interface consists of a function which the
configuration shall provide to invoke the language processor.

\emph{We could do \texttt{REQUIRES} in a macro-expansion way by adding
an argument to \texttt{Config\_SourceChar} to specify the source file.
However, I'm assuming we will prefer to recursively ``run'' each
required file. One of the results of that will be the classes and
methods made public by that \texttt{REQUIRES} subject.}

@@

\lstinputlisting[language=rexx,label=rexxapistart.rexx,caption=rexxapistart.rexx]{rexxapistart.rexx}

\begin{itemize}
\tightlist
\item
  \texttt{How} is one of
  \texttt{\textquotesingle{}COMMAND\textquotesingle{}},
  \texttt{\textquotesingle{}FUNCTION\textquotesingle{}}, or
  \texttt{\textquotesingle{}SUBROUTINE\textquotesingle{}} and indicates
  how the program is invoked.
\end{itemize}

\emph{What does OOI say for How when running REQUIREd files?}

\begin{itemize}
\item
  \texttt{Source} is an identification of the source of the program to
  be processed.
\item
  \texttt{Environment} is the initial value of the environment to be
  used in processing commands. This has components for the name of the
  environment and how the input and output of commands is to be
  directed.
\item
  \texttt{Arguments} is the initial argument list to be used in
  processing. This has components to specify the number of arguments,
  which arguments are omitted, and the values of arguments that are not
  omitted.
\item
  \texttt{Streams} has components for the default input stream to be
  used and the default output streams to be used.
\item
  \texttt{Traps} is the list of traps to be used in processing (see
  nnn). This has components to specify whether each trap is omitted or
  not.
\end{itemize}

\subsubsection{Semantics:}\label{semantics}

@@

If the program was terminated due to a \texttt{RETURN} or \texttt{EXIT}
instruction without an expression the completion response is
\texttt{\textquotesingle{}N\textquotesingle{}}.

If the program was terminated due to a \texttt{RETURN} or \texttt{EXIT}
instruction with an expression the indicator in the completion response
is \texttt{\textquotesingle{}R\textquotesingle{}} and the description of
the completion response is the value of the expression.

If the program was terminated due to an error the indicator in the
completion response is \texttt{\textquotesingle{}E\textquotesingle{}}
and the description in the completion response comprises information
about the error that terminated processing.

If How was \texttt{\textquotesingle{}REQUIRED\textquotesingle{}} and the
completion response was not
\texttt{\textquotesingle{}E\textquotesingle{}}, the \texttt{Provides}
argument is set to reference classes made available. See nnn for the
semantics of these classes.

@@

\begin{itemize}
\tightlist
\item
  \emph{syntactic\_characters};
\item
  \emph{extra\_letters};
\item
  \emph{other\_blank\_characters};
\item
  \emph{other\_negators};
\item
  \emph{other\_characters}.
\end{itemize}

@@

The following characters represent the category of characters called
\emph{syntactic\_characters}, identified by their names. The glyphs used
to represent them in this document are also shown.
\emph{Syntactic\_characters} shall be available in every configuration:

\begin{itemize}
\item
  \& ampersand;
\item
  ' apostrophe, single quotation mark, single quote;
\item
  \begin{itemize}
  \tightlist
  \item
    asterisk, star;
  \end{itemize}
\item
  blank, space;
\item
  A-Z capital letters A through Z;
\item
  : colon;
\item
  , comma;
\item
  0-9 digits zero through nine;
\item
  = equal sign;
\item
  ! exclamation point, exclamation mark;
\item
  \textgreater{} greater-than sign;
\item
  \begin{itemize}
  \tightlist
  \item
    hyphen, minus sign;
  \end{itemize}
\item
  \textless{} less-than sign;
\item
  {[} left bracket, left square bracket;
\item
  ( left parenthesis;
\item
  \% percent sign;
\item
  . period, decimal point, full stop, dot;
\item
  \begin{itemize}
  \tightlist
  \item
    plus sign;
  \end{itemize}
\item
  ? question mark;
\item
  '' quotation mark, double quote;
\item
  ~reverse slant, reverse solidus, backslash;
\item
  {]} right bracket, right square bracket;
\item
  ) right parenthesis;
\item
  ; semicolon;
\item
  / slant, solidus, slash;
\item
  a-z small letters a through z;
\item
  \textasciitilde{} tilde, twiddle;
\item
  \_ underline, low line, underscore;
\item
  vertical line, bar, vertical bar.
\end{itemize}

@@

A configuration may have a category of characters in source programs
called \emph{extra\_letters}. \emph{Extra\_letters} are determined by
the configuration.

@@

A configuration may have a category of characters in source programs
called \emph{other\_blank\_characters}. \emph{Other\_blank\_characters}
are determined by the configuration. Only the following characters
represent possible characters of this category:

\begin{itemize}
\tightlist
\item
  carriage return;
\item
  form feed;
\item
  horizontal tabulation;
\item
  new line;
\item
  vertical tabulation.
\end{itemize}

@@

A configuration may have a category of characters in source programs
called \emph{other\_negators}. \emph{Other\_negators} are determined by
the configuration. Only the following characters represent possible
characters of this category. The glyphs used to represent them in this
document are also shown:

\begin{itemize}
\tightlist
\item
  \^{} circumflex accent, caret;
\item
  ¬ not sign.
\end{itemize}

@@

A configuration may have a category of characters in source programs
called \emph{other\_characters}. \emph{Other\_characters} are determined
by the configuration.

@@

The configuration characters and encoding interface consists of
functions which the configuration shall provide which are concerned with
the encoding of characters.

\begin{itemize}
\tightlist
\item
  \texttt{Config\_SourceChar};
\item
  \texttt{Config\_OtherBlankCharacters};
\item
  \texttt{Config\_Upper};
\item
  \texttt{Config\_Compare};
\item
  \texttt{Config\_B2C};
\item
  \texttt{Config\_C2B};
\item
  \texttt{Config\_Substr};
\item
  \texttt{Config\_Length};
\item
  \texttt{Config\_Xrange}.
\end{itemize}

@@

\subsubsection{Syntax:}\label{syntax}

\lstinputlisting[language=rexx,label=configsourcechar.rexx,caption=configsourcechar.rexx]{configsourcechar.rexx}

\subsubsection{Semantics:}\label{semantics-1}

Supply the characters of the source program in sequence, together with
the \emph{EOL} and \emph{EOS} events. The \emph{EOL} event represents
the end of a line. The \emph{EOS} event represents the end of the source
program. The \emph{EOS} event must only occur immediately after an
\emph{EOL} event. Either a character or an event is supplied on each
invocation, by setting \texttt{\#Outcome}.

If this function is unable to supply a character because the source
program encoding is incorrect the indicator of the completion response
is \texttt{\textquotesingle{}E\textquotesingle{}} and the description of
the completion information is the encoding which is incorrect, in
hexadecimal notation.

@@

\subsubsection{Syntax:}\label{syntax-1}

\lstinputlisting[language=rexx,label=configotherblankcharacters.rexx,caption=configotherblankcharacters.rexx]{configotherblankcharacters.rexx}

\subsubsection{Semantics:}\label{semantics-2}

Get \emph{other\_blank\_characters} (see nnn).

Set \texttt{\#Outcome} to a string of zero or more distinct characters
in arbitrary order. Each character is one that the configuration
considers equivalent to the character \emph{Blank} for the purposes of
parsing.

@@

\subsubsection{Syntax:}\label{syntax-2}

Config\_Upper(Character)

Character is the character to be translated to uppercase.

\subsubsection{Semantics:}\label{semantics-3}

Translate Character to uppercase. Set \#Outcome to the translated
character. Characters which have been subject to this translation are
referred to as being in uppercase. Config\_Upper applied to a character
in uppercase must not change the character.

@@

\subsubsection{Syntax:}\label{syntax-3}

Config\_Lower(Character)

Character is the character to be translated to lowercase.

\subsubsection{Semantics:}\label{semantics-4}

Translate Character to lowercase. Set \#Outcome to the translated
character. Characters which have been subject to this translation are
referred to as being in lowercase. Config\_Lower applied to a character
in lowercase must not change the character. Config\_Upper of the outcome
of Config\_Lower(Character) shall be the original character.

@@

\subsubsection{Syntax:}\label{syntax-4}

Config Compare(Characterl, Character2)

@@

Character1 is the character to be compared with Character2.

Character2 is the character to be compared with Character1.

\subsubsection{Semantics:}\label{semantics-5}

@@

\begin{itemize}
\item
  `equal' if Character1 is equal to Character2;
\item
  `greater' if Character1 is greater than Character2;
\item
  `lesser' if Character' is less than Character2.
\item
  The function shall exhibit the following characteristics. If Config
  \_Compare(a,b) produces
\item
  `equal' then Config\_Compare(b,a) produces `equal';
\item
  `greater' then Config\_Compare(b,a) produces `lesser';
\item
  `lesser' then Config\_Compare(b,a) produces `greater';
\item
  `equal' and Config\_Compare(b,c) produces `equal' then
  Config\_Compare(a,c) produces `equal';
\item
  `greater' and Config\_Compare(b,c) produces `greater' then Config
  \_Compare(a,c) produces `greater';
\item
  `lesser' and Config \_Compare(b,c) produces `lesser' then
  Config\_Compare(a,c) produces `lesser';
\item
  `equal' then Config\_Compare(a,c) and Config\_Compare(b,c) produce the
  same value.
\item
  Syntactic characters which are different characters shall not compare
  equal by Config\_Compare, see nnn.
\end{itemize}

@@

\subsubsection{Syntax:}\label{syntax-5}

Config B2C (Binary)

Binary is a sequence of digits, each `0' or `1'. The number of digits
shall be a multiple of eight.

\subsubsection{Semantics:}\label{semantics-6}

@@

\subsubsection{Syntax:}\label{syntax-6}

Config C2B (String)

String is a string.

\subsubsection{Semantics:}\label{semantics-7}

Translate String to a sequence of digits, each `0' or `1'. Set \#Outcome
to the result. This function is the inverse of Config\_B2C.

@@

\subsubsection{Syntax:}\label{syntax-7}

Config Substr(String, n)

String is a string.

n is an integer identifying a position within String.

\subsubsection{Semantics:}\label{semantics-8}

Copy the n-th character from String. The leftmost character is the first
character. Set Outcome to the resulting character.

If this function is unable to supply a character because there is no
n-th character in String the indicator of the completion response is
`M'.

If this function is unable to supply a character because the encoding of
String is incorrect the indicator of the completion response is `E' and
the description of the completion information is the encoding which is
incorrect, in hexadecimal notation.

@@

\subsubsection{Syntax:}\label{syntax-8}

Config Length (String)

String is a string.

\subsubsection{Semantics:}\label{semantics-9}

Set \#Outcome to the length of the string, that is, the number of
characters in the string.

If this function is unable to determine a length because the encoding of
String is incorrect, the indicator of the completion response is `E' and
the description of the completion information is the encoding which is
incorrect, in hexadecimal notation.

@@

\subsubsection{Syntax:}\label{syntax-9}

Config Xrange(Characterl, Character2)

Character1 is the null string, or a single character.

Character2 is the null string, or a single character.

\subsubsection{Semantics:}\label{semantics-10}

If Character1 is the null string then let LowBound be a lowest ranked
character in the character set according to the ranking order provided
by Config\_Compare; otherwise let LowBound be Character1.

If Character2 is the null string then let HighBound be a highest ranked
character in the character set according to the ranking order provided
by Config\_Compare; otherwise let HighBound be Character2lf \#Outcome
after Config\_Compare(LowBound,HighBound) has a value of

@@

\subsubsection{Syntax:}\label{syntax-10}

Config ObjectNew

\subsubsection{Semantics:}\label{semantics-11}

@@

\subsubsection{Syntax:}\label{syntax-11}

Config Array Size(Object, size)

Object is an object.

Size is an integer greater or equal to 0.

\subsubsection{Semantics:}\label{semantics-12}

The configuration should prepare to deal efficiently with the object as
an array with indexes having values up to the value of size.

@@

\subsubsection{Syntax:}\label{syntax-12}

Config Array Put(Array, Item, Index)

Array is an array.

Item is an object

Index is an integer greater or equal to 1.

\subsubsection{Semantics:}\label{semantics-13}

The configuration shall record that the array has Item associated with
Index.

@@

\subsubsection{Syntax:}\label{syntax-13}

Config Array At(Array, Index)

Array is an array.

Index is an integer greater or equal to 1.

\subsubsection{Semantics:}\label{semantics-14}

The configuration shall return the item that the array has associated
with Index.

@@

\subsubsection{Syntax:}\label{syntax-14}

Config Array At(Array, Index)

Array is an array.

Index is an integer greater or equal to 1.

\subsubsection{Semantics:}\label{semantics-15}

Return `1' if there is an item in Array associated with Index, `0'
otherwise.

@@

\subsubsection{Syntax:}\label{syntax-15}

Config Array At(Array, Index)

Array is an array.

Index is an integer greater or equal to 1.

\subsubsection{Semantics:}\label{semantics-16}

After this operation, no item is associated with the Index in the Array.

@@

\subsubsection{Syntax:}\label{syntax-16}

Config Command(Environment, Command)

Environment is the environment to be addressed. It has components for:

\begin{itemize}
\item
  the name of the environment;
\item
  the name of a stream from which the command will read its input. The
  null string indicates use of the default input stream;
\item
  the name of a stream onto which the command will write its output. The
  null string indicates use of the default output stream. There is an
  indication of whether writing is to APPEND or REPLACE;
\item
  the name of a stream onto which the command will write its error
  output. The null string indicates use of the default error output
  stream. There is an indication of whether writing is to APPEND or
  REPLACE.
\end{itemize}

Command is the command to be executed.

\subsubsection{Semantics:}\label{semantics-17}

Perform a command.

\begin{itemize}
\item
  set the indicator to `E' or `F' if the command ended with an ERROR
  condition, or a FAILURE condition, respectively;
\item
  set \#RC to the return code string of the command.
\end{itemize}

@@

The external routines interface consists of a function which the
configuration shall provide to invoke external routines.

See nnn and nnn for a description of the language features that use
external routines.

@@

\subsubsection{Syntax:}\label{syntax-17}

Config ExternalRoutine(How, NameType, Name, Environment, Arguments,
Streams, Traps) where:

How is one of `FUNCTION' or `SUBROUTINE' and indicates how the external
routine is to be invoked.

@@

\subsubsection{Semantics:}\label{semantics-18}

@@

OOI has external classes explicitly via the ::CLASS abc EXTERNAL
mechanism. Analogy with classic would also allow the subject of
::REQUIRES to be coded in non-Rexx. However ::REQUIRES subject is coded,
we need to gather in knowledge of its method names because of the search
algorithm that determines which method is called. Hence reasonable that
the ultimate external call is to a method. Perhaps combine
Config\_ExternalRoutine with Config\_ExternalMethod.

There is a terminology clash on ``environment''. Perhaps easiest to
change the classic to ``address\_environment''. (And make it part of new
``environment''?)

There are terminology decisions to make about ``files'', ``programs'',
and ``packages''. Possibly ``program'' is the thing you run (and we
don't say what it means physically), ``file'' is a unit of scope
(ROUTINEs in current file before those in REQUIREd), and ``package'' we
don't use (since a software package from a shop would probably have
several files but not everything to run a program.) Using ``file'' this
way may not be too bad since we used ``stream'' rather than ``tile'' in
the classic definition.

The How parameter will need `METHOD' as a value. Should API\_Start also
allow 'METHOD''. If we pass the new Environment we don't have to pass
Streams separately.

Text of Config\_ExternalMethod waiting on such decisions.

\subsubsection{Syntax:}\label{syntax-18}

Config ExternalMethod (How, NameType, Name, Environment, Arguments,
Streams, Traps)

How is one of `FUNCTION' or `SUBROUTINE' and indicates how the external
routine is to be invoked.

NameType is a specification of whether the name was provided as a symbol
or as a string literal.

Name is the name of the routine to be invoked.

Arguments is a specification of the arguments to the routine, with the
same components as on API\_Start.

\subsubsection{Semantics:}\label{semantics-19}

\section{External data queue}\label{external-data-queue}

The external data queue interface consists of functions which the
configuration shall provide to manipulate an external data queue
mechanism.

See nnn, nnn, nnn, nnn, and nnn for a description of language features
that use the external data queue. The configuration shall provide an
external data queue mechanism. The following functions shall be
provided:

@@

\subsubsection{Syntax:}\label{syntax-19}

Config Push(String)

String is the value to be retained in the external data queue.

\subsubsection{Semantics:}\label{semantics-20}

Add String as an item to the end of the external data queue from which
Config\_Pull will remove an item.

@@

\subsubsection{Syntax:}\label{syntax-20}

Config Queue (String)

String is the value to be retained in the external data queue.

\subsubsection{Semantics:}\label{semantics-21}

Add String as an item to the opposite end of the external data queue
from which Config\_Pull will remove an item.

@@

\subsubsection{Syntax:}\label{syntax-21}

Config Pull()

\subsubsection{Semantics:}\label{semantics-22}

Retrieve an item from the end of the external data queue to which
Config\_Push adds an element to the list. Set \#Outcome to the value of
the retrieved item.

If no item could be retrieved the indicator of the completion response
is `F'.

@@

\subsubsection{Syntax:}\label{syntax-22}

Config Queued ()

\subsubsection{Semantics:}\label{semantics-23}

@@

\subsubsection{Syntax:}\label{syntax-29}

Config Stream Unique ()

\subsubsection{Semantics:}\label{semantics-30}
